---
title: "Homework 5"
author: "Joyce"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: spacelab
    keep_md: yes
---
need to look at the readme file to answer the questions
## Instructions
Answer the following questions and/or complete the exercises in RMarkdown. Please embed all of your code and push the final work to your repository. Your report should be organized, clean, and run free from errors. Remember, you must remove the `#` for any included code chunks to run.  

## Load the tidyverse
```{r message=FALSE, warning=FALSE}
library("tidyverse")
library("janitor")
```

## Data 
For this assignment, we will use data from a study on vertebrate community composition and impacts from defaunation in [Gabon, Africa](https://en.wikipedia.org/wiki/Gabon). One thing to notice is that the data include 24 separate transects. Each transect represents a path through different forest management areas.  

Reference: Koerner SE, Poulsen JR, Blanchard EJ, Okouyi J, Clark CJ. Vertebrate community composition and diversity declines along a defaunation gradient radiating from rural villages in Gabon. _Journal of Applied Ecology_. 2016. This paper, along with a description of the variables is included inside the data folder.   

**1. Load `IvindoData_DryadVersion.csv` and store it as a new object called `gabon`.**
```{r}
gabon <- read_csv("data/IvindoData_DryadVersion.csv")
```

**2. Use one or more of the summary functions you have learned to get an idea of the structure of the data.**  
```{r}
summary(gabon)
```

```{r}
glimpse(gabon)
```
  
**3. Use `mutate()` Change the variables `HuntCat`, `LandUse`, and `TransectID` to factors.**
```{r}
gabon <- gabon %>% 
  mutate(HuntCat = as.factor(HuntCat)) %>% 
  mutate(LandUse = as.factor(LandUse)) %>% 
  mutate(TransectID = as.factor(TransectID))

is.factor(gabon$HuntCat)
is.factor(gabon$LandUse)
is.factor(gabon$TransectID)
```



**4. Use `filter` to make three new dataframes focused only on 1. national parks, 2. logging concessions, and 3. neither. Have a look at the README in the data folder so you understand the variables.**
```{r}
gabon_2 <- filter(gabon, LandUse == "Park")
gabon_3 <- filter(gabon, LandUse == "Logging")
gabon_4 <- filter(gabon, LandUse == "Neither")
```

**5. How many transects are recorded for each land use type?**
```{r}
gabon_2
gabon_3
gabon_4
```

National parks has 7 transect ID, which 2 of the 7 are in the same park, Logging has 13 transect ID, and neither has 4 transect ID.
**6. For which land use type (national parks, logging, or neither) is average all species diversity the greatest?**
```{r}

```

```{r}

```

```{r}

```

**7. Use `filter` to find the transect that has the highest relative abundance of elephants. What land use type is this? Use `arrange()` to sort your results.** 
```{r}

```

**8. Use `filter` to find all transects that have greater than 15 tree species or a breast height diameter between 50 and 60cm.  **
```{r}

```

**9.Which transects and land use types have more than 10 tree species and 10 mammal species? Use `arrange()` to sort by the number of tree species.**
```{r}

```

**10. Explore the data! Develop one question on your own that includes at least two lines of code. **
```{r}

```

## Knit and Upload
Please knit your work as a .pdf or .html file and upload to Canvas. Homework is due before the start of the next lab. No late work is accepted. Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  